---
title: "data_prism_naive_models_2"
author: "Samantha M. Roth"
date: "May 21, 2020"
output: html_document
---

Load some packages that might be relevant and the data.
```{r}
load("Data_prism")
library(car)
library(glmnet)
library(openintro)
# library(interactions)
#library(sjPlot)
library(ggplot2)
```

```{r}
hist(Data_prism1$Yield);  hist(Data_prism1$GDD); hist(Data_prism1$EDD);hist(sqrt(Data_prism1$EDD)); hist(Data_prism1$Pr)
```
The histogram for sqrt(EDD) looks much less skewed than the histogram for EDD.
```{r}
sqrtEDD<- sqrt(Data_prism1$EDD)
Data_prism1<- as.data.frame(cbind(Data_prism1, sqrtEDD))
```
We should first look at the pairs plot
```{r}
pairs(~Yield+GDD+sqrtEDD+Pr+year, data=Data_prism1)
```
Hmm it looks like GDD and sqrtEDD are almost linearly related.

```{r}
fit= lm(Yield~ GDD + sqrtEDD + Pr + GDD:sqrtEDD + GDD:Pr + sqrtEDD:Pr, data= Data_prism1)
summary(fit)
AIC(fit)
vif(fit)
outlierTest(fit)
```
The VIFs are very large for all of the regressors. This is bad. So a simple linear regression model worked much bettter for Yield_anomaly, GDD_anomaly, EDD_anomaly, and SPI than for the raw data. 
