---
title: "exploring_spatial_dependencies2"
author: "Samantha M. Roth"
date: "June 16, 2020"
output: html_document
---

```{r}
setwd("c:/Users/smr31/advising related/climate and crop research- summer 2020/Climate_CornYield-master-new/Climate_CornYield-master")
library(sp)
library(maps)
library(maptools)
library(ncdf4)
library(ggplot2)
library(usmap)
library(precintcon)
library(BMS)
library(adaptMCMC)
library(geoR) #for variogram
library(housingData) #for returning county centroid
library(gstat)
library(fields)
library(binaryLogic)
load("Data_prism")
```

```{r}
locations<- cbind(Data_prism$lat, Data_prism$lon)
```
Create the variogram
```{r}
vg_aic<- variog(coords= locations, data= Data_prism$residual_aicfit)
plot(vg_aic)
```

```{r}
#vfit_aic<- variofit(vario= vg_aic, ini.cov.pars= c(1,1))
#summary(vfit_aic)
init_s2<- seq(from= 1, to= 1000, length= 1000)
init_phi<- seq(from= .01, to= 100, length= length(init_s2))
init<- cbind(init_s2, init_phi)
vfit_aic<- variofit(vario= vg_aic, ini.cov.pars = init)
summary(vfit_aic)
```
From testing different sets of initial values to mimize the loss function, the estimates are simga^2= 241.004497 and phi= 9.247984.

Next, we'll plot the variogram for aicfit.
```{r}
geomodel1<- likfit(coords = locations, data = Data_prism$Yield,
trend = "cte", ini.cov.pars= c(241.004497,9.247984))

```


we want to fit this model: Yield~ GDD + EDD + poly(Pr, 2) + as.numeric(year) + Tmax + Tmin + 
    lastSF + firstFF + GDD:EDD + GDD:poly(Pr, 2) + poly(Pr, 2):EDD + 
    GDD:Tmax + GDD:Tmin + GDD:lastSF + GDD:firstFF + EDD:Tmax + 
    EDD:Tmin + EDD:lastSF + EDD:firstFF + poly(Pr, 2):Tmax + 
    poly(Pr, 2):Tmin + poly(Pr, 2):lastSF + poly(Pr, 2):firstFF + 
    Tmax:Tmin + Tmax:lastSF + Tmax:firstFF + Tmin:lastSF + Tmin:firstFF + 
    lastSF:firstFF
