---
title: "exploring spatial dependencies of residuals"
author: "Samantha M. Roth"
date: "June 5, 2020"
output: html_document
---

Load data and necessary packages. 
Residual plot for each year for the model fit2 in the file variable_selection_nofips_yearasint, slightly altered from Haochen's code. Need to have first created a model and added the residuals for that model to the dataframe Data_prism.
```{r}
setwd("c:/Users/smr31/advising related/climate and crop research- summer 2020/Climate_CornYield-master-new/Climate_CornYield-master")
library(sp)
library(maps)
library(maptools)
library(ncdf4)
library(ggplot2)
library(usmap)
library(precintcon)
library(BMS)
library(adaptMCMC)
library(geoR) #for variogram
library(housingData) #for returning county centroid
library(gstat)
library(fields)
library(binaryLogic)
load("Data_prism")
#Data_prism<-Data_prism[complete.cases(Data_prism), ]
#Data_prism$StateANSI<-factor(Data_prism$StateANSI)
#Data_prism$year<-as.numeric(Data_prism$year)

dir.create("fig_res_aicfit")
#load("variable_selection_nofips_yearasint")
for (i in 1:32){
  filename<-paste("fig_res_aicfit/aicfit",i,".jpeg",sep="")
  jpeg(file = filename,width = 800,height=800)
  index<-which(Data_prism$year==i)
  a=plot_usmap(regions = "counties",include=c("PA","NY","NJ","MD","DE","DC","NC","VA","SC","WV","OH","MI","GA","KY","IN","IL","AL","TN","WI","MS","MN","MO","LA","AR","IA"),
               data=Data_prism[index,], values = "residual_aicfit") + labs(title = paste("Yield residual of each county in ",i+1980,sep=""))+ 
    scale_fill_gradient2(low = "blue",mid="white", high ="red", limits=c(-45,45),name="bu/acre")+theme(plot.title = element_text(size=14))
  plot(a)
  dev.off()
}
```


Residual plot for each year, slightly altered from Haochen's code. Need to have first created a model and added the residuals for that model to the dataframe Data_prism.
```{r}
setwd("c:/Users/smr31/advising related/climate and crop research- summer 2020/Climate_CornYield-master-new/Climate_CornYield-master")

dir.create("fig_res_fit2")
for (i in 1:32){
  filename<-paste("fig_res_fit2/fit2",i,".jpeg",sep="")
  jpeg(file = filename,width = 800,height=800)
  index<-which(Data_prism$year==i)
  a=plot_usmap(regions = "counties",include=c("PA","NY","NJ","MD","DE","DC","NC","VA","SC","WV","OH","MI","GA","KY","IN","IL","AL","TN","WI","MS","MN","MO","LA","AR","IA"),
               data=Data_prism[index,], values = "residual_fit2") + labs(title = paste("Yield residual of each county in ",i+1980,sep=""))+ 
    scale_fill_gradient2(low = "blue",mid="white", high ="red", limits=c(-45,45),name="bu/acre")+theme(plot.title = element_text(size=14))
  plot(a)
  dev.off()
}
```
