---
title: "spatial modeling- subsetted data"
author: "Samantha M. Roth"
date: "6/18/2020"
output: html_document
---

```{r}
setwd("c:/Users/smr31/advising related/climate and crop research- summer 2020/Climate_CornYield-master-new/Climate_CornYield-master")
library(sp)
library(maps)
library(maptools)
library(ncdf4)
library(ggplot2)
library(usmap)
library(precintcon)
library(BMS)
library(adaptMCMC)
library(geoR) #for variogram
library(housingData) #for returning county centroid
library(gstat)
library(fields)
library(binaryLogic)
load("Data_prism")
```
Define the subset of locations of interest (in our first example it is just year 1, 1981)
```{r}
loc_sub<- cbind(Data_prism$lat[which(Data_prism$year==1)], Data_prism$lon[which(Data_prism$year==1)])
data_sub<- Data_prism$Yield[which(Data_prism$year==1)]
fips_sub<- Data_prism$fips[which(Data_prism$year==1)]
```

We subset the data to only include the first year of data and turn it into a geodata object. 
```{r}
prism_sub<- Data_prism[Data_prism$year==1,]
gd_prism_sub<- as.geodata(prism_sub, coords.col= 14:15, data.col = 1, covar.col = c("GDD", "EDD", "Pr", "Tmax", "Tmin", "lastSF", "firstFF"))
summary(gd_prism_sub)
```
Create the variogram
```{r}
vg<- variog(geodata= gd_prism_sub)
plot(vg)
```

```{r}
#vfit_aic<- variofit(vario= vg_aic, ini.cov.pars= c(1,1))
#summary(vfit_aic)
init_s2<- seq(from= 1, to= 1000, length= 1000)
init_phi<- seq(from= .01, to= 100, length= length(init_s2))
init<- cbind(init_s2, init_phi)
vfit<- variofit(vario= vg, ini.cov.pars = init)
summary(vfit)
```
From testing different sets of initial values to minimize the loss function, the estimates are simga^2= 960.9062 and phi= 5.4799.

```{r}
trend1<- trend.spatial(trend= "cte", geodata= gd_prism_sub, add.to.trend = ~ GDD + EDD + Pr + Tmax + Tmin + lastSF + firstFF)

trend2<-  trend.spatial(trend= "cte", geodata= gd_prism_sub, add.to.trend = ~ GDD + EDD + Pr + I(Pr^2) + Tmax + Tmin + lastSF + firstFF)

trend3<- trend.spatial(trend="cte", geodata = gd_prism_sub, add.to.trend= ~  GDD + EDD + Pr + I(Pr^2) + Tmax + Tmin + lastSF + firstFF+ I(GDD*EDD) + I(GDD*Pr) + I(Pr *EDD) + I(GDD*Tmax) + I(GDD*Tmin) + I(GDD*lastSF) + I(GDD*firstFF) + I(EDD*Tmax) + I(EDD*Tmin) + I(EDD*lastSF) + I(EDD*firstFF) + I(Pr*Tmax) + I(Pr*Tmin) + I(Pr*lastSF) + I(Pr*firstFF) + I(Tmax*Tmin) + I(Tmax*lastSF) + I(Tmax*firstFF) + I(Tmin*lastSF) + I(Tmin*firstFF) + I(lastSF*firstFF))
```

Next, we'll train the model to predict Yield with GDD, EDD, Pr, Tmax, Tmin, lastSF, firstFF, and 
```{r}
geomodel1<- likfit(geodata= gd_prism_sub, trend= trend1, ini.cov.pars= c(960.9062, 5.4799))
summary(geomodel1)
resids<- resid(geomodel1)
```
```{r}
geomodel2<- likfit(geodata= gd_prism_sub, trend= trend2, ini.cov.pars= c(960.9062, 5.4799))
summary(geomodel2)
resids<- resid(geomodel2)
```

```{r}
geomodel3<- likfit(geodata= gd_prism_sub, trend= trend3, ini.cov.pars= c(960.9062, 5.4799))
summary(geomodel3)
resids<- resid(geomodel3)
```

Create a dataframe for the residuals
```{r}
resid_data<- as.data.frame(matrix(rep(NA),nrow= length(resids), ncol=1))
resid_data$residual<- resids
resid_data$fips<- fips_sub
resid_data$lat<- loc_sub[,1]
resid_data$lon<- loc_sub[,2]
head(resid_data)
```
Plot the residuals over space.
```{r}
library(sp)
library(maps)
library(maptools)
library(ncdf4)
library(ggplot2)
library(usmap)
library(precintcon)
library(BMS)
library(adaptMCMC)
library(geoR) #for variogram
library(housingData) #for returning county centroid
library(gstat)
library(fields)
library(binaryLogic)
#year 1's res plot
dir.create("fig_res_spatial2")
filename<-paste("fig_res_spatial2/",1,".jpeg",sep="")
jpeg(file = filename,width = 800,height=800)
  a=plot_usmap(regions = "counties", include=c("PA","NY","NJ","MD","DE","DC","NC","VA","SC","WV","OH","MI","GA","KY","IN","IL","AL","TN","WI","MS","MN","MO","LA","AR","IA"), data=resid_data, values = "residual") + labs(title = paste("Yield residual of each county in ",1981,sep=""))+ scale_fill_gradient2(low = "blue",mid="white", high ="red", limits=c(-20,20), name="bu/acre")+theme(plot.title = element_text(size=14))
  plot(a)
  dev.off()
```



