---
title: "variable_selection_nofips_yearfactor"
author: "Samantha M. Roth"
date: "June 11, 2020"
output: html_document
---

Load data, packages.
```{r}
setwd("c:/Users/smr31/advising related/climate and crop research- summer 2020/Climate_CornYield-master-new/Climate_CornYield-master")
library(sp); library(maps); library(maptools); library(ncdf4); library(ggplot2); library(usmap); library(precintcon); library(BMS); library(adaptMCMC); library(geoR) #for variogram
library(housingData) #for returning county centroid
library(gstat); library(fields); library(binaryLogic); library(car)
load("Data_prism")
Data_prism<-Data_prism[complete.cases(Data_prism), ]
Data_prism$StateANSI<-factor(Data_prism$StateANSI)
Data_prism$year<-factor(Data_prism$year)
#Data_prism$sqrtEDD<- sqrt(Data_prism$EDD)
#Data_prism$sqrtPr<- sqrt(Data_prism$Pr)
```


Full model with all interactions:
```{r}
fit<- lm(Yield~ GDD+EDD+ poly(Pr,2) + year + Tmax + Tmin + lastSF + firstFF + GDD:EDD + GDD:poly(Pr,2) + poly(Pr,2):EDD + GDD:Tmax +GDD:Tmin + GDD:lastSF + GDD:firstFF + EDD:Tmax + EDD:Tmin + EDD:lastSF + EDD:firstFF + poly(Pr,2):Tmax + poly(Pr,2):Tmin + poly(Pr,2):lastSF + poly(Pr,2):firstFF + Tmax:Tmin + Tmax:lastSF + Tmax:firstFF + Tmin:lastSF +Tmin:firstFF + lastSF:firstFF, data= Data_prism) 
summary(fit); AIC(fit); vif(fit)
```
Full Model:
Residual standard error: 22.12 on 47431 degrees of freedom
Multiple R-squared:  0.5789,	Adjusted R-squared:  0.5783 
F-statistic: 987.9 on 66 and 47431 DF,  p-value: < 2.2e-16
AIC: 429028.7
NPE: 0.2084848

Ten-fold cross validation
```{r}
tfcv_fit<- function(d= Data_prism){
  totlen=dim(d)[1]
  cvind<-c(1:totlen)
  groups<-split(cvind,sample(rep(1:10,each = floor(totlen/10))))
  CV_each<-rep(NA,10)
  sse<- rep(NA,10)
  for (k in 1:10) {  #10 cross-validation tests
    test_indx<-groups[[k]]
    test_data<-d[test_indx, ]
    train_indx<-cvind[-test_indx]
    train_data<- d[train_indx,]
    train_model<- lm(Yield ~ GDD+EDD+ poly(Pr,2) + year + Tmax + Tmin + lastSF + firstFF + GDD:EDD + GDD:poly(Pr,2) + poly(Pr,2):EDD + GDD:Tmax +GDD:Tmin + GDD:lastSF + GDD:firstFF + EDD:Tmax + EDD:Tmin + EDD:lastSF + EDD:firstFF + poly(Pr,2):Tmax + poly(Pr,2):Tmin + poly(Pr,2):lastSF + poly(Pr,2):firstFF + Tmax:Tmin + Tmax:firstFF + Tmin:lastSF +Tmin:firstFF + lastSF:firstFF, data= train_data)
    pred<-predict(train_model,test_data) #which variables to test?
    CV_each[k]<-mean((pred - test_data$Yield)^2)
    sse[k]<- sum((pred - test_data$Yield)^2)
  }
  NPE<-sqrt(mean(CV_each))/mean(d[ ,"Yield"],na.rm=TRUE)
  SSPE<- sum(sse)
  return(list("NPE" = NPE, "SSPE"=SSPE))
}
tfcv_fit()
```

Next we'll use stepwise AIC variable selection to see if we get any improvements
```{r}
aicfit= step(fit)
summary(aicfit)
AIC(aicfit)
```
aicfit: 
lm(formula = Yield ~ GDD + EDD + poly(Pr, 2) + year + Tmax + 
    Tmin + lastSF + firstFF + GDD:EDD + GDD:poly(Pr, 2) + EDD:poly(Pr, 
    2) + GDD:Tmax + GDD:Tmin + GDD:lastSF + GDD:firstFF + EDD:Tmax + 
    EDD:Tmin + EDD:lastSF + EDD:firstFF + poly(Pr, 2):Tmax + 
    poly(Pr, 2):Tmin + poly(Pr, 2):lastSF + poly(Pr, 2):firstFF + 
    Tmax:Tmin + Tmax:firstFF + Tmin:lastSF + Tmin:firstFF + lastSF:firstFF, 
    data = Data_prism)
Residual standard error: 22.12 on 47432 degrees of freedom
Multiple R-squared:  0.5789,	Adjusted R-squared:  0.5783 
F-statistic:  1003 on 65 and 47432 DF,  p-value: < 2.2e-16
AIC: 429026.8
NPE: 0.2084889

```{r}
fitCGS<- lm(Yield~ GDD_changingGS +EDD_changingGS + poly(Pr_changingGS,2) + year + Tmax + Tmin + lastSF + firstFF + GDD_changingGS:EDD_changingGS + GDD_changingGS:poly(Pr_changingGS,2) + poly(Pr_changingGS,2):EDD_changingGS + GDD_changingGS:Tmax + GDD_changingGS:Tmin + GDD_changingGS:lastSF + GDD_changingGS:firstFF + EDD_changingGS:Tmax + EDD_changingGS:Tmin  + EDD_changingGS:lastSF +EDD_changingGS:firstFF + poly(Pr_changingGS,2):Tmax + poly(Pr_changingGS,2):Tmin + poly(Pr_changingGS,2):lastSF + poly(Pr_changingGS,2):firstFF + Tmax:Tmin +Tmax:lastSF + Tmax:firstFF + Tmin:lastSF +Tmin:firstFF + lastSF:firstFF, data= Data_prism) 
summary(fitCGS); AIC(fitCGS); vif(fitCGS)
```
fitCGS: worse
Residual standard error: 23.23 on 47431 degrees of freedom
Multiple R-squared:  0.5357,	Adjusted R-squared:  0.5351 
F-statistic: 829.3 on 66 and 47431 DF,  p-value: < 2.2e-16
AIC: 433661.5
NPE: 0.218913

```{r}
fit2<- lm(Yield~ GDD+EDD+ poly(Pr,2) + as.factor(year) + Tmax + Tmin + lastSF + firstFF + GDD:EDD + GDD:poly(Pr,2) + poly(Pr,2):EDD + GDD:Tmax +GDD:Tmin + GDD:lastSF + GDD:firstFF + EDD:Tmax + EDD:Tmin + EDD:lastSF + EDD:firstFF + poly(Pr,2):Tmax + poly(Pr,2):Tmin + poly(Pr,2):lastSF + poly(Pr,2):firstFF + Tmax:Tmin + Tmax:firstFF + Tmin:lastSF +Tmin:firstFF + lastSF:firstFF, data= Data_prism) 
summary(fit2); AIC(fit2); vif(fit2)
Data_prism$hindcast_fit2<- predict(fit2, Data_prism)
Data_prism$residual_fit2<-Data_prism$Yield-Data_prism$hindcast_fit2 #the residuals
```
BEST MODEL:
fit2: no Tmax:lastSF
Residual standard error: 22.12 on 47432 degrees of freedom
Multiple R-squared:  0.5789,	Adjusted R-squared:  0.5783 
F-statistic:  1003 on 65 and 47432 DF,  p-value: < 2.2e-16
AIC: 429026.8
NPE: 0.2084734

```{r}
beta.hat=fit2$coef #get the coefficient estimates
yhat=fitted(fit2)
eps=resid(fit2)
plot(yhat,eps,main="y-hat vs residuals")
abline(h=0,col="red",lwd=2)
qqnorm(eps)
qqline(eps)
```
The residual plot looks pretty good, it seems as though there are many more medium values of yhat that small and large values of yhat. This is backed up by the histogram below.
```{r}
hist(yhat)
```
The appearance of possible fanning is potentially due to the distribution of yhat.
```{r}
source("C:/Users/smr31/STAT511/partial residual plots.r")
## mean structure
prp(fit2,data= Data_prism,names=c("GDD"))
prp(fit2,data= Data_prism,names=c("EDD"))
prp(fit2,data= Data_prism,names=c("Pr"))
prp(fit2,data= Data_prism,names=c("year"))
prp(fit2,data= Data_prism,names=c("lastSF"))
prp(fit2,data= Data_prism,names=c("firstFF"))
prp(fit2,data= Data_prism,names=c("Tmax"))
prp(fit2,data= Data_prism,names=c("Tmin"))
```